---
title: "Reproducible Finance with R"
author: "jonathan.regenstein@rstudio.com"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
install.packages("tidyverse")
install.packages("tidyquant")
install.packages("timetk")
install.packages("tibbletime")
install.packages("broom")
install.packages("dygraphs")

devtools::install_github("jbkunst/highcharter")
```


```{r}
library(tidyverse)
library(tidyquant)
library(tibbletime)
library(timetk)
library(dygraphs)
library(highcharter)
library(broom)
```

### Introducing R

    + Statistical programming language -> by data scientists, for data scientists
    + Base R + 17,000 packages
    + RStudio
    + Shiny
    + sparklyr
    + tensorflow
    + Rmarkdown
    + database connectors
    + htmlwidgets
  


### Packages for finance


```{r, eval = FALSE}
library(PerformanceAnalytics)
library(PortfolioAnalytics)
library(TTR)
library(tidyquant)
library(quantmod)
library(xts)
```


List of packages for finance here:
https://cran.r-project.org/web/views/Finance.html

### Packages for data visualization


```{r, eval = FALSE}
library(ggplot2)
library(dygraphs)
library(highcharter)
library(shiny)
```


### An example project

- Import data for 5 ETFs
- Visualize prices and returns
- Calculate some stats of interest
- Create an SMA 50 v. SMA 200
- Run a rolling linear model, chart some results

    + SPY (S&P500 fund)
    + EFA (a non-US equities fund) 
    + IJS (a small-cap value fund) 
    + EEM (an emerging-mkts fund)
    + AGG (a bond fund)

### Import the Data


    + from excel using `read_excel`
    
    + from csv using `read_csv`
   
    + from Yahoo! Finance using `getSymbols` or `tq_get`
 
    + from myssql, postgres etc using `dbConnect`

    + via API with `httr` and `jsonlite` (important for Alternative Data)
  
 
### Get data

```{r}

# The symbols vector holds our tickers. 
symbols <- c("SPY",
             "EFA", 
             "IJS", 
             "EEM",
             "AGG")

# data imported from Yahoo! Finance
etf_prices<-
  tq_get(symbols, from  = "2012-12-31") %>% 
  select(symbol, date, adjusted) %>% 
  spread(symbol, adjusted) %>% 
  tk_xts(date_var = date)
```

### Inspect the data


```{r}
head(etf_prices)
```

### Visualize

```{r}
dygraph(etf_prices)
```

### Another Visualization

```{r}
highchart(type = "stock") %>% 
  hc_add_series(etf_prices[,1]) %>% 
  hc_add_series(etf_prices[,2]) %>% 
  hc_add_series(etf_prices[,3]) %>% 
  hc_add_series(etf_prices[,4]) %>% 
  hc_add_series(etf_prices[,5]) %>% 
  hc_title(text = "Highcharting 5 ETFS") %>%
  hc_yAxis(opposite = FALSE,
           labels = list(format = "${value}")) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_navigator(enabled = FALSE) %>% 
  hc_exporting(enabled = TRUE)
```


### Performance Analytics: getting started


```{r}
etf_returns <- 
  Return.calculate(etf_prices, method = "log")

head(etf_returns)
```


```{r}
table.Stats(etf_returns)
```


```{r}
table.CAPM(etf_returns, etf_returns$SPY)
```

### Scatter returns

```{r}
etf_returns %>% 
  tk_tbl(rename_index = "date") %>% 
  select(-date) %>% 
  gather(symbol, return, -SPY) %>% 
  ggplot(aes(x = SPY, y = return, color = symbol)) +
  geom_point(alpha = .5) +
  facet_wrap(~symbol)

```

### Add regression line

```{r}
etf_returns %>% 
  tk_tbl(rename_index = "date") %>% 
  dplyr::select(-date) %>% 
  gather(symbol, return, -SPY) %>% 
  ggplot(aes(x = SPY, y = return, color = symbol)) +
  geom_point(alpha = .5) + 
  geom_smooth(formula = y ~ x, se = TRUE) +
  facet_wrap(~symbol)
```


### Interactive Scatter

```{r}
etf_returns %>% 
  tk_tbl(rename_index = "date") %>% 
  hchart(., type = "scatter", hcaes(x = SPY, y = EEM, date = date)) %>% 
  hc_xAxis(title = list(text = "Market Returns"),
           labels = list(format = "{value}%")) %>% 
  hc_yAxis(title = list(text = "EEM Returns"),
           labels = list(format = "{value}%")) %>% 
  hc_title(text = "Emerging Market v. SPY") %>% 
  hc_tooltip(pointFormat = "date: {point.date} <br> 
             EEM return: {point.y:.4f}  <br> 
             mkt return: {point.x:.4f}")
```


Grab beta or slope of regression line.

```{r}
slope <- table.CAPM(etf_returns, etf_returns$SPY)[2, 2]
```

Add the regression line to the original scatter

```{r}

etf_returns_tibble <- 
  etf_returns %>% 
  tk_tbl(rename_index = "date")

  hchart(etf_returns_tibble, type = "scatter", 
         hcaes(x = SPY, y = EEM, date = date)) %>%
  hc_add_series(etf_returns_tibble, "line", 
                hcaes(x = SPY, 
                      y = SPY * slope)) %>% 
  hc_xAxis(title = list(text = "Market Returns"),
           labels = list(format = "{value}%")) %>% 
  hc_yAxis(title = list(text = "EEM Returns"),
           labels = list(format = "{value}%")) %>% 
  hc_title(text = "Scatter with Beta Line")
```

### Other Nice Functions (too many to list)

```{r}
table.DownsideRisk(etf_returns, Rf= .0003)
```

```{r}
table.Drawdowns(etf_returns$EEM)
```

```{r}
SharpeRatio(etf_returns, Rf = .0003)
```

```{r}
InformationRatio(etf_returns, etf_returns$SPY)
```

```{r}
SemiDeviation(etf_returns)
```


### Standard Deviation of each asset

```{r}
StdDev(na.omit(etf_returns))
```

### Portfolio Standard Deviation

```{r}
StdDev(na.omit(etf_returns), 
       weights = c(.1, .2, .2, .2, .3))
```


### Contribution to Portfolio Standard Deviation  

```{r}
StdDev(na.omit(etf_returns), 
       weights = c(.1, .2, .2, .2, .3), 
       portfolio_method = "component")
```

### Visualize Contribution

```{r}
StdDev(na.omit(etf_returns), 
       weights = c(.1, .2, .2, .2, .3), 
       portfolio_method = "component") %>% 
  as.tibble() %>% 
  add_column(symbols = sort(symbols)) %>% 
  ggplot(aes(x = symbols, y = pct_contrib_StdDev, fill = symbols)) +
  geom_col(width = .5) +
  labs(y = "percent contribution to vol", x = "", title = "Asset Contribution to Vol") +
  scale_y_continuous(labels = scales::percent)
```


### Rolling Mean Calculations and Visualization: our own functions

```{r}
sma_50 <- rollify(mean, window = 50)
sma_200 <- rollify(mean, window = 200)
sd_50 <- rollify(sd, window = 50)
sd_200 <- rollify(sd, window = 200)
 
etf_rolling_calculations <-  
  etf_prices %>% 
  tk_tbl(rename_index = "date") %>% 
  select(date, SPY) %>% 
  mutate(sma50 = sma_50(SPY),
         sma200 = sma_200(SPY),
         sd200_lower = sma200 - sd_200(SPY),
         sd200_upper = sma200 + sd_200(SPY),
         signal = ifelse(sma50 > sma200, 1, 0)
         ) %>%
  select(-SPY) %>% 
  na.omit()
```

```{r}
highchart()%>%
  hc_add_series(etf_rolling_calculations, type = "line", 
                hcaes(x = date, y = sma200), 
                name = "sma200"
                ) %>%
 hc_add_series(etf_rolling_calculations, type = "line", 
                hcaes(x = date, y = sma50), 
                name = "sma50",
               color = "green") %>%
  hc_add_series(etf_rolling_calculations, 
                type = "arearange", 
                hcaes(x = date, 
                      low = sd200_lower, 
                      high = sd200_upper),
                color = "pink",
                alpha = .25,
                fillOpacity = 0.3,
                showInLegend = FALSE
                ) %>%
  hc_title(text = "SMA 50 v. SMA 200") %>% 
      hc_xAxis(type = 'datetime') %>% 
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_exporting(enabled = TRUE) 
```

### Calculate growth of dollar

SMA trend strategy versus buy and hold

```{r}
etf_trend <- 
  etf_prices %>% 
  tk_tbl(rename_index = "date") %>% 
  select(date, SPY) %>% 
  mutate(sma50 = sma_50(SPY),
         sma200 = sma_200(SPY),
         returns = log(SPY) - log(lag(SPY)),
         signal = ifelse(sma50 > sma200, 1, 0),
         daily_treas = (1 + (2/100)) ^ (1/252) - 1,
         buy_hold_returns = (.9 * returns) + (.1 * daily_treas),
         trend_returns = if_else(lag(signal) == 1, (signal * returns), daily_treas)
         ) %>%
  na.omit() %>% 
  mutate(
         trend_growth = accumulate(1 + trend_returns, `*`),
         buy_hold_growth = accumulate(1 + buy_hold_returns, `*`)) %>%
  select(date, trend_growth, buy_hold_growth) %>% 
  tk_xts(date_var = date) 
```

### Visualize Strategy versus Buy and Hold

```{r}

  
  highchart(type = "stock") %>% 
  hc_title(text = "Growth") %>%
  hc_add_series(etf_trend$trend_growth, color = "cornflowerblue",  name = "trend") %>%
  hc_add_series(etf_trend$buy_hold_growth, color = "green", name = "buy_hold") %>%
  hc_add_theme(hc_theme_flat()) %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE) %>% 
  hc_legend(enabled = TRUE)
```

### Rolling Models

```{r}

rolling_lm <- rollify(.f = function(EEM, SPY) {
                              lm(EEM ~ SPY)
                           }, 
                      window = 100, 
                      unlist = FALSE)

etf_returns_tibble %>%
  select(date, EEM, SPY) %>%
  na.omit() %>% 
  mutate(rolling_beta = rolling_lm(EEM, SPY)) %>% 
  slice(-1:-99) %>% 
  mutate(glanced = map(rolling_beta, 
                      glance)) %>% 
  unnest(glanced)

```

```{r}
rolling_model_results <- 
  etf_returns_tibble %>%
  select(date, EEM, SPY) %>%
  na.omit() %>% 
  mutate(rolling_beta = rolling_lm(EEM, SPY)) %>% 
  slice(-1:-99) %>% 
  mutate(glanced = map(rolling_beta, 
                      glance)) %>% 
  unnest(glanced) %>% 
  select(date, r.squared, adj.r.squared, p.value)

rolling_model_results %>% 
  hchart(., hcaes(x = date, y = r.squared), type = "line")  %>% 
  hc_title(text = "Rolling R-Squared")
  
```


### Other Packages of Interest

```{r, eval = FALSE}
library(forecast) # Good out of the box forecasting tools. Useful for macro trends.
library(h2o) # machine learning libraries
library(keras) # deep learning tensorflow.rstudio.com
library(lime) # for ML white-boxing
library(ranger) # random forest
library(recipes) # for ML preprocessing
library(rsample) # for resampling
library(caret) # classification and regression
library(tidytext) # parse text and mining
library(tidyposterior) # posthoc after resampling
```

### Learn more
www.reproduciblefinance.com

http://r4ds.had.co.nz/

### Get Started

- Download R: https://cloud.r-project.org/

- Download RStudio: www.rstudio.com/products/rstudio/download/#download

- datacamp course: www.datacamp.com/tracks/applied-finance-with-r


