---
title: "ETF Thoughts"
always_allow_html: yes
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

Start with a Shiny app: 
- choose an asset, choose what to display, show charts and tables below

Rex's notes
- emphasize reproducibility and ease of data import
- show from excel and from yahoo? why both? if use a public resource, it's fully reproducible

- show the performance analytics packaage as way to talk about the package ecosystem, tidyquant
- emphasize that R can do a lot with a little code
-  https://cran.r-project.org/web/views/Finance.html
- visualization is a great feature of R, let's look at that data and how simply can create 
- why is R growing, why is it popular and to me the right tool
- some ML thoughts? 

- R is by stats for stats, simpler syntax. Not only does that make it faster to pick up and learn, it leads to readable code, and that leads to reproducibility. 

```{r setup, include = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
library(tidyverse)
library(tidyquant)
library(tibbletime)
library(timetk)
library(dygraphs)
library(highcharter)
library(writexl)
library(readxl)

sma_50 <- rollify(mean, window = 50)
sma_200 <- rollify(mean, window = 200)
sd_50 <- rollify(sd, window = 50)
sd_200 <- rollify(sd, window = 200)

```

### Packages for finance

PerformanceAnalytics
PortfolioAnalytics
TTR
tidyquant
quantmod
xts

### Packages more general

ggplot2 - most popular data vis
dygraph - interactive data vis
highcharter - interactive data vis
billboarder  interactive data vis
leaflet - map buiilding
Rmarkdown - create reports
Shiny - create interactive dashboards 

### Learn More

`R for Data Science`
`www.reproduciblefinance.com` - book coming soon, email me for a copy

### An example project

Import data for 5 ETFs, visualize prices and returns, calculate some stats or interest, create an SMA 50 v. SMA 200 chart.

We will work with price and returns data for 5 ETFS:

    + SPY (S&P500 fund)
    + EFA (a non-US equities fund) 
    + IJS (a small-cap value fund) 
    + EEM (an emerging-mkts fund)
    + AGG (a bond fund)

### Import the Data

Three ways to get our data: 
- Excel using `read_excel`
- yahoo! Finance using `getSymbols` or`tq_get`
- sql, or some internal database, using `dbConnect`

We'll go with yahoo! Finance to make this reproducible. [why does that matter?]



```{r}

# The symbols vector holds our tickers. 
symbols <- c("SPY",
             "EFA", 
             "IJS", 
             "EEM",
             "AGG")

# etf_prices <- 
#     getSymbols(symbols, 
#                src = 'yahoo',
#                auto.assign = TRUE, 
#                warnings = FALSE) %>% 
#     map(~Ad(get(.))) %>% 
#     reduce(merge) %>%
#     `colnames<-`(symbols)

etf_prices<-
  tq_get(symbols, from  = "2012-12-31") %>% 
  select(symbol, date, adjusted) %>% 
  spread(symbol, adjusted) %>% 
  tk_xts(date_var = date)


head(etf_prices)
```
 

```{r, eval = FALSE}
etf_prices <- 
 read_excel("etf-prices.xlsx") %>%
  mutate(date = ymd(date)) %>% 
  tk_xts(date_var = date)
```


```{r}
plot(etf_prices)
```

```{r}
dygraph(etf_prices)
```

```{r}
highchart(type = "stock") %>% 
  hc_add_series(etf_prices[,1]) %>% 
  hc_add_series(etf_prices[,2]) %>% 
  hc_add_series(etf_prices[,3]) %>% 
  hc_add_series(etf_prices[,4]) %>% 
  hc_add_series(etf_prices[,5])
```


Let's calculate returns using the `Return.calculate` function, which will process the columns as vectors.


```{r}
etf_returns <- 
  Return.calculate(etf_prices, method = "log")

head(etf_returns)
```

```{r}
plot(etf_returns)
```


```{r}
dygraph(etf_returns, main = "ETF Returns") %>%
  dyRangeSelector()
```

Scatter EEM returns against SPY returns

```{r}
etf_returns %>% 
  tk_tbl(rename_index = "date") %>% 
  hchart(., type = "scatter", hcaes(x = SPY, y = EEM, date = date)) %>% 
  hc_xAxis(title = list(text = "Market Returns")) %>% 
  hc_yAxis(title = list(text = "EEM Returns")) %>% 
  hc_title(text = "Emerging Market v. SPY") %>% 
  hc_tooltip(pointFormat = "date: {point.date} <br> 
             EEM return: {point.y:.4f}  <br> 
             mkt return: {point.x:.4f}")
```


```{r}
etf_returns %>% 
  tk_tbl(rename_index = "date") %>% 
  select(-date) %>% 
  gather(symbol, return, -SPY) %>% 
  ggplot(aes(x = SPY, y = return, color = symbol)) +
    geom_point(alpha = .5) + 
  geom_smooth(formula = y ~ x, se = TRUE) +
  facet_wrap(~symbol)
```


PerformanceAnalytics

```{r}
table.Stats(etf_returns)
```


```{r}
table.CAPM(etf_returns, etf_returns$SPY, Rf = .0003)
```


```{r}
table.DownsideRisk(etf_returns, Rf= .0003)
```


```{r}
table.Drawdowns(etf_returns$IJS)
```


```{r}
SharpeRatio(etf_returns, Rf = .0003)
```


```{r}
InformationRatio(etf_returns, etf_returns$SPY)
```


```{r}
SemiDeviation(etf_returns)
```


```{r}
chart.RelativePerformance(etf_returns, etf_returns$SPY, main = "Relative Performance to S&P")
```



```{r}
chart.RollingPerformance(etf_returns, 
		FUN = 'SharpeRatio.annualized', 
		width = 100,
		main = "Rolling 100-day Sharpe Ratio")
```


One more...create a portfolio and calculate contribution to standard deviation. 

```{r}
StdDev(na.omit(etf_returns), 
       weights = c(.1, .2, .2, .2, .3), 
       portfolio_method = "component") %>% 
  as.tibble() %>% 
  add_column(symbols = sort(symbols)) %>% 
  ggplot(aes(x = symbols, y = pct_contrib_StdDev, fill = symbols)) +
  geom_col(width = .5) +
  labs(y = "percent contribution to vol", x = "", title = "Asset Contribution to Vol") +
  scale_y_continuous(labels = scales::percent)
```

A more involved look at our returns. Let's calculate rolling 50 day and 200 day means, and rolling 200 day standard deviation, and chart them to see when the 50 days is consistently outside that range. [JKR: we have no limitations except our own creativity, and we could wrap this whole thing into a function or a package.]


```{r}
 
etf_rolling_calculations <-  
  etf_prices %>% 
  tk_tbl(rename_index = "date") %>% 
  select(date, SPY) %>% 
  mutate(sma50 = sma_50(SPY),
         sma200 = sma_200(SPY),
         sd200_lower = sma200 - sd_200(SPY),
         sd200_upper = sma200 + sd_200(SPY)) %>% 
  select(-SPY) %>% 
  tk_xts(date_var = date)

dygraph(etf_rolling_calculations, main = "SMA 50 v. SMA 200") %>% 
  dySeries(c("sd200_lower", "sma200", "sd200_upper"), label = "sma200") %>%
  dyShading(from = "2013-09-01", to = "2014-03-01", 
            axis = "x", 
            color = "#ffe6e6") %>% 
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 2))  %>% 
  dyCrosshair(direction = "vertical") %>%
  dyRangeSelector()
```




Learn More: 

R 4 data Science - http://r4ds.had.co.nz/
Reproducible Finance with R - http://www.reproduciblefinance.com/ (email me for an advanced copy)

Other topics of interest

`forecast` package. Good out of the box forecasting tools. Useful for macro trends.
`h20` machine learning libraries. 
`keras` deep learning. see `tensorflow.rstudio.com`
`lime` for ML white-boxing
